Local Spectral Clustering of Density Upper Level Sets

### 2019-05-23 neurips/local_spectral_density_clustering_neurips.tex
Through intro, saved a bit of space

\PassOptionsToPackage{numbers}{natbib}
\usepackage{neurips_2019}

@@

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

@@

\theoremstyle{aldenrmrk}
\newtheorem{remark}{Remark}

@@

\begin{document}

\maketitle

\begin{abstract}
\vskip 0.1 in
  Spectral clustering methods are a family of popular nonparametric clustering
  tools.  Recent works have proposed and analyzed \emph{local} spectral methods, such as Personalized PageRank (\ppr),
  which extract clusters using locally-biased random walks around a user-specified
  seed node.  In contrast to existing results, we analyze PPR in a traditional
  statistical learning setup, where we obtain samples from an unknown
  distribution, and aim to identify connected regions of high-density (density
  clusters).  We prove that PPR, run on a neighborhood graph, extracts
  sufficiently salient density clusters, and provide empirical support for our theory.
\end{abstract}

@@

Let $\Xbf = \{x_1, \ldots, x_n\}$ be a sample drawn i.i.d.\ from a
distribution $\Pbb$ on $\Rd$, with density $f$, and consider the problem of 
clustering: splitting the data into groups which satisfy some notion of
within-group similarity and between-group difference.  We focus on spectral
clustering methods, a family of powerful nonparametric clustering algorithms.
Roughly speaking, a spectral technique first constructs a geometric graph $G$,
where vertices are associated with samples, and edges correspond to proximities
between samples. It then learns a feature embedding based on the Laplacian of
$G$, and applies a simple clustering technique (such as k-means clustering) in
the embedded feature space.

When applied to geometric graphs constructed from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to recent increased interest in
local spectral algorithms, which leverage locally-biased spectra computed using
random walks around a user-specified seed node.  A popular local clustering
algorithm is Personalized PageRank (\ppr), first introduced by
\citet{haveliwala2003}, and further developed in
\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013},
among others.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, which has led
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (like
conductance).  In this paper, we adopt a more traditional statistical viewpoint,
and examine what the output of a local clustering algorithm on $\Xbf$ reveals
about the unknown density $f$.  In particular, we examine the ability of the \pprspace algorithm to recover \emph{density clusters} of $f$, which are defined as the
connected components of the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$
for some threshold $\lambda > 0$ (a central object of interest in the
classical statistical literature on clustering, dating back to
\citet{hartigan1981}).

\paragraph{PPR on a neighborhood graph} We now describe the clustering algorithm that will be our focus for the rest of 
the paper. We start with the geometric graph that we form based on the samples 
$\Xbf$: for a radius $r > 0$, we consider the \emph{$r$-neighborhood graph} of 
$\Xbf$, denoted $G_{n,r}=(V,E)$, an unweighted, undirected graph with vertices
$V=\Xbf$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i - x_j}
\leq  r$, where $\norm{\cdot}$ denotes Euclidean norm. We denote by $\Abf \in \Reals^{n \times n}$ the adjacency matrix, with entries $\Abf_{uv} = 1$ if $(u,v) \in E$ and $0$ otherwise, by $\Dbf$ the diagonal degree matrix, with $\Dbf_{uu} = \sum_{v \in V} \Abf_{uv}$, and by $\Ibf{}$ the $n \times n$ identity matrix.

Next, we define the PPR vector $\pbf = \pbf(v,\alpha;G_{n,r})$, with respect to  
a seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, to be
the solution of the following linear system:
\begin{equation}
\label{eqn: ppr_vector}
\pbf = \alpha \ebf{v} + (1 - \alpha) \pbf \Wbf,
\end{equation}
where $\Wbf = (\Ibf{} + \Dbf^{-1}\Abf)/2$ is the lazy random walk matrix over $G_{n,r}$ 
and $e_{v}$ denotes the indicator vector for node $v$ (with a 1 in the $v$th
position and 0 elsewhere).

For a level $\beta > 0$ and a target volume $\vol_0 > 0$, we define a
\emph{$\beta$-sweep cut} of $\pbf = (p_u)_{u \in V}$ as  
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta = \set{u \in V: \frac{p_u}{\Dbf_{uu}} > \frac{\beta}{\vol_{0}}}.
\end{equation}
We will use normalized cut to determine which sweep cut $S_{\beta}$ is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V \setminus S$, we define the cut as $\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c} \Abf_{uv}$, the volume as $\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$, and the \emph{normalized cut} as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{11})$},\footnote{The choice of a specific range such as 
\smash{$(\frac{1}{40}, \frac{1}{11})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}, we output the cluster estimate $\widehat{C} = S_{\beta^*}$ which has minimum normalized cut $\Phi(S_{\beta^{\star}}; G_{n,r})$. For concreteness, we summarize this procedure in Algorithm \ref{alg: ppr}.

\begin{algorithm}
	\caption{PPR on a Neighborhood Graph}
	\label{alg: ppr}	
	{\bfseries Input:} data $\Xbf=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation 
	parameter $\alpha \in [0,1]$, seed $v \in \Xbf$, target stationary volume $\vol_0 >
	0$. \\   
	{\bfseries Output:} cluster $\Cest \subseteq V$.
	\begin{algorithmic}[1]
		\STATE Form the neighborhood graph $G_{n,r}$.
		\STATE Compute the PPR vector $\pbf(v, \alpha; G_{n,r})$ as in \eqref{eqn:
			ppr_vector}. 
		\STATE For $\beta \in (\frac{1}{40}, \frac{1}{11})$ compute sweep cuts
		$S_{\beta}$ as in \eqref{eqn: sweep_cuts}.
		\STATE Return \smash{$\Cest = S_{\beta^*}$}, where 
		$$
		\beta^* = \argmin_{\beta \in (\frac{1}{40}, \frac{1}{11})} \Phi(S_{\beta}; G_{n,r}).
		$$
	\end{algorithmic}
\end{algorithm}

\paragraph{Estimation of density clusters}

Let $\Cbb_f(\lambda)$ denote the connected components of the density upper level
set $\{x \in \Rd: f(x) > \lambda\}$.  For a given density cluster $\Cset \in
\Cbb_f(\lambda)$, we call $\Cset[\Xbf] = \Cset \cap \Xbf$ the \emph{empirical
density cluster}. The symmetric set difference between estimated and empirical cluster is perhaps the most frequently used metric to quantify cluster estimation error \citep{korostelev2012,polonik1995,rigollet2009}.

\begin{definition}[Symmetric set difference]
	\label{def: symmetric_set_diff}
	For an estimator \smash{$\Cest \subseteq \Xbf$} and set
	$\mathcal{S} \subseteq \Reals^d$, the symmetric set difference of $\Cest$ and $\mathcal{S} \cap \Xbf = \mathcal{S}[\Xbf]$ is 
	\begin{equation}
	\label{eqn: misclassification_rate}
	\Delta(\Cest, \mathcal{S}) := \abs{\Cest \setminus \mathcal{S}[\Xbf] \cup \mathcal{S}[\Xbf] \setminus \Cest}.
	\end{equation}
\end{definition}

However, the symmetric set difference does not account for the distance points in $\widehat{C} \setminus \mathcal{S}[\Xbf]$ may be from $\mathcal{S}$ \citep{singh2009}. We therefore give a second notion of cluster estimation, first introduced by \citet{hartigan1981} and defined asymptotically, which measures whether $\widehat{C}$ can distinguish any two distinct elements $\Cset, \Cset' \in \Cbb_f(\lambda)$.

\begin{definition}[Consistent density cluster estimation]
	\label{def: consistent_density_cluster_estimation}
	For an estimator \smash{$\Cest \subseteq \Xbf$} and cluster 
	$\Cset \in \Cbb_f(\lambda)$, we say \smash{$\Cest$} is a consistent
	estimator of $\Cset$ if for all $\Cset' \in \Cbb_f(\lambda)$ with $\Cset \not=
	\Cset'$ the following holds as $n \to \infty$: 
	\begin{equation}
	\label{eqn: consistent_density_cluster_recovery}
	\Cset[\Xbf] \subseteq \Cest \quad \text{and} \quad
	\Cest \cap \Cset'[\Xbf] = \emptyset,
	\end{equation}
	with probability tending to 1.
\end{definition}
\paragraph{Summary of results}
A summary of our main results (and outline for the rest of this paper) is as
follows.

\begin{enumerate}
	\item In Section \ref{sec: consistent_cluster_estimation_with_ppr}, we introduce a set of natural geometric conditions on the density cluster $\Cset$, formalize a measure of difficulty based on these geometric conditions, and show that when Algorithm \ref{alg: ppr} is properly initialized, the symmetric set difference of $\widehat{C}$ and a thickened version of the density cluster $\Csig$ is upper bounded by this difficulty measure.

	\item We further show that if the density cluster $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr} will consistently estimate a density cluster in the sense of \eqref{eqn: consistent_density_cluster_recovery}.

	\item In Section \ref{sec: analysis}, we detail some of the analysis required to prove our main results, and expose the part various geometric quantities play in the ultimate difficulty of the clustering problem.

	\item In Section \ref{sec: experiments}, we empirically
	demonstrate the tightness of our analysis, and provide examples showing how violations of the geometric conditions we require manifestly
	impact density cluster recovery by \ppr.  
\end{enumerate}

Our main takeaway can be summarized as follows: \ppr, run on a neighborhood graph, recovers geometrically compact high-density clusters.

\paragraph{Related work}

In addition to the background given previously, a few related lines of work are worth
highlighting. Similar in spirit to our results are the works
\citep{shi2009,schiebinger2015}, who study the consistency of
spectral algorithms in recovering the latent labels in certain
nonparametric mixture models. These results focus on global rather than local
algorithms, and as such impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density
clusters, which is the focus in our work. Perhaps most importantly, these works rely on general cluster saliency conditions, which implicitly depend on many distinct geometric aspects of the cluster $\Cset$ under consideration. We make this dependence more explicit, and in doing so expose the role each geometric condition plays in the clustering problem.

Additionally, we note that density clustering and level set estimation is a well-studied problem. \citep{polonik1995, rigollet2009} study density clustering under symmetric set difference, \citep{tsybakov1997, singh2009} exhibit minimax optimal level set estimators under Hausdorff loss and \citep{hartigan1981, chaudhuri2010} consider consistent estimation of the cluster tree, to note but a few works on the subject. Our goal is not to improve on these results, or offer yet another algorithm for level set estimation; indeed, seen as a density clustering algorithm, \pprspace has none of the optimality guarantees of the previous works. Rather, our motivation is to better understand and characterize the distinctions between those density clusters which are well conditioned for local spectral algorithms, and those which are not.

@@

To motivate these conditions, for an arbitrary graph $G = (V,E)$ and subset of vertices $S \subseteq V$, consider the normalized cut $\Phi(S;G)$ (defined as in \eqref{eqn: normalized_cut}), as well as the mixing time of a random walk over the induced subgraph $G[S]$ (defined in Section \ref{sec: analysis} by \eqref{eqn: mixing_time}, and denoted by $\tau_{\infty}(G[S])$). In \citet{zhu2013}, it is shown that for a constant $c > 0$, the \pprspace estimate $\widehat{C}$ of $S$ satisfies
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\widehat{C} \setminus S; G) + \vol(S \setminus \widehat{C}; G) \leq c \bigl( \Phi(S,G) \cdot \tau_{\infty}(G[S])\bigr) \vol(S;G)
\end{equation}
where the left hand side resembles a (degree-weighted) form of the symmetric set difference of \eqref{eqn: misclassification_rate}.

@@

\begin{theorem}
	\label{thm: misclassification_rate}
	Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a $\kappa$-well conditioned density cluster (with respect to some $\theta$), and additionally assume $f$ satisfies \ref{asmp: valid_region}. Then, there exists a universal constant $c_4 > 0$ such that with probability tending to $1$ as $n \to \infty$, 
	\begin{equation}
	\label{eqn: misclassification_rate_ub}
	\Delta(\Csig[\Xbf], \Cest) \leq c_4 \kappa(\Cset) \frac{\Lambda_{\sigma}}{\lambda_{\min}}.
	\end{equation}
\end{theorem}

@@

\begin{theorem}
	\label{thm: consistent_recovery_of_density_clusters}
	Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a $\kappa$-well conditioned cluster (with respect to some $\theta$), and additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm \ref{alg: ppr} is well-initialized, there exists a universal constant $c_5 > 0$ such that if
	\begin{equation}
	\label{eqn: kappa_ub}
	\kappa(\Cset) \leq c_5 \frac{\lambda_{\sigma}^2r^d \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)},
	\end{equation}
	then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}.
\end{theorem}

>>>>

%\PassOptionsToPackage{numbers}{natbib}
\usepackage{neurips_2019}

@@

%%%%%% Begin Alden
\DeclarePairedDelimiter\abs{\lvert}{\rvert}

@@

\theoremstyle{aldenrmrk}
\newtheorem{remark}{Remark}
%%%%%% End Alden

@@

\begin{document}
\maketitle

\begin{abstract}
We analyze Personalized PageRank (PPR), a local spectral method for clustering,
which extract clusters using locally-biased random walks around a user-specified
seed node.  In constrast to pervious work, we adopt a traditional statistical
learning setup, where we obtain samples from an unknown distribution, and aim to
identify connected regions of high-density (density clusters).  We prove that
PPR, run on a neighborhood graph, extracts sufficiently salient density
clusters. We also provide empirical support for our theory.
\end{abstract}

@@

In this paper, we study the problem of clustering: splitting a given data set
into groups that satisfy some notion of within-group similarity and
between-group difference.  We focus on spectral clustering, a family of powerful
nonparametric clustering algorithms.  Generally speaking, a spectral algorithm
first constructs a geometric graph $G$, where vertices correspond to samples,
and edges correspond to proximities between samples. It then learns a feature
embedding based on the Laplacian of $G$, and applies a simple clustering
technique (like k-means clustering) in the embedded feature space.

When applied to geometric graphs built from a large number of samples,
global spectral clustering methods can be computationally cumbersome and   
insensitive to the local geometry of the underlying distribution
\citep{leskovec2010,mahoney2012}.  This has led to recent increased interest in
local spectral algorithms, which leverage locally-biased spectra computed using
random walks around a user-specified seed node.  A popular local clustering
algorithm is Personalized PageRank (PPR), first introduced by
\citet{haveliwala2003}, and then further developed by
\citep{spielman2011,spielman2014,andersen2006,mahoney2012,zhu2013},
among others.

Local spectral clustering techniques have been practically very successful
\citep{leskovec2010,andersen2012,gleich2012,mahoney2012,wu2012}, leading 
many authors to develop supporting theory
\citep{spielman2013,andersen2009,gharan2012,zhu2013} that gives worst-case
guarantees on traditional graph-theoretic notions of cluster quality (such as
conductance).  In this paper, we adopt a more traditional statistical viewpoint,
and examine what the output of local clustering on a data set reveals about the
underlying density $f$.  In particular, we examine the ability of PPR to recover
\emph{density clusters} of $f$, defined as the connected components of
the upper level set $\{x \in \Rd : f(x) \geq \lambda\}$ for some $\lambda > 0$
(a central object of interest in the statistical clustering literature, dating
back to \citet{hartigan1981}).

\paragraph{PPR on a neighborhood graph.} We now describe the clustering
algorithm that will be our focus for the rest of the paper. Let $\Xbf = \{x_1,
\ldots, x_n\}$ be a sample drawn i.i.d.\ from a distribution $\Pbb$ on $\Rd$,
with density $f$.  For a radius $r > 0$, we define $G_{n,r}=(V,E)$ to be the
\emph{$r$-neighborhood graph} of $\Xbf$, an unweighted, undirected graph with
vertices $V=\Xbf$, and an edge $(x_i,x_j) \in E$ if and only if $\norm{x_i -
x_j} \leq r$, where $\norm{\cdot}$ is the $\ell_2$ norm. We denote by $\Abf \in
\Reals^{n \times n}$ the adjacency matrix, with entries $\Abf_{uv} = 1$ if
$(u,v) \in E$ and $0$ otherwise.  We also denote by $\Dbf$ the diagonal degree
matrix, with $\Dbf_{uu} = \sum_{v \in V} \Abf_{uv}$, and by $\Ibf{}$ the $n
\times n$ identity matrix.

Next, we define the PPR vector $\pbf = \pbf(v,\alpha;G_{n,r})$, based on
a seed node $v \in V$ and a teleportation parameter $\alpha \in [0,1]$, to be 
the solution of the following linear system:
\begin{equation}
\label{eqn: ppr_vector}
\pbf = \alpha \ebf{v} + (1 - \alpha) \pbf \Wbf,
\end{equation}
where $\Wbf = (\Ibf{} + \Dbf^{-1}\Abf)/2$ is the lazy random walk matrix over
$G_{n,r}$ and $e_{v}$ is the indicator vector for node $v$ (that has a 1 in the
$v$th position and 0 elsewhere). For a level $\beta > 0$ and a target volume
$\vol_0 > 0$, we define a \emph{$\beta$-sweep cut} of $\pbf = (p_u)_{u \in V}$
as   
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta := \set{u \in V: \frac{p_u}{\Dbf_{uu}} > \frac{\beta}{\vol_{0}}}.
\end{equation}
We will use the normalized cut metric to determine which sweep cut $S_{\beta}$
is the best cluster estimate. For a set $S \subseteq V$ with complement $S^c = V
\setminus S$, we define \smash{$\cut(S;G_{n,r}) := \sum_{u \in S, v \in S^c}
\Abf_{uv}$}, and \smash{$\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$}.  We
define the \emph{normalized cut} of $S$ as
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{11})$}\footnote{The choice of a specific range such as  
\smash{$(\frac{1}{40}, \frac{1}{11})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}, we output the cluster estimate
\smash{$\Cest = S_{\beta^*}$} that has minimum normalized cut.
%\smash{$\Phi(S_{\beta^*}; G_{n,r})$}. 
For concreteness, this is summarized in Algorithm \ref{alg: ppr}.

\begin{algorithm}
\caption{PPR on a neighborhood graph}
\label{alg: ppr}	
{\bfseries Input:} data $\Xbf=\{x_1,\ldots,x_n\}$, radius $r > 0$, teleportation
parameter $\alpha \in [0,1]$, seed $v \in \Xbf$, target stationary volume
$\vol_0 > 0$. \\     
{\bfseries Output:} cluster $\Cest \subseteq V$.
\begin{algorithmic}[1]
  \STATE Form the neighborhood graph $G_{n,r}$.
  \STATE Compute the PPR vector $p=\pbf(v, \alpha; G_{n,r})$ as in \eqref{eqn: 
    ppr_vector}. 
  \STATE For \smash{$\beta \in (\frac{1}{40}, \frac{1}{11})$} compute sweep cuts 
  $S_{\beta}$ as in \eqref{eqn: sweep_cuts}.
  \STATE Return as a cluster \smash{$\Cest = S_{\beta^*}$}, where  
  $$
  \beta^* = \argmin_{\beta \in (\frac{1}{40}, \frac{1}{11})} \Phi(S_{\beta}; G_{n,r}).
  $$
\end{algorithmic}
\end{algorithm}

\paragraph{Estimation of density clusters} Let \smash{$\Cbb_f(\lambda)$} denote
the connected components of the density upper level set $\{x \in \Rd: f(x) >
\lambda\}$.  For a given density cluster \smash{$\Cset \in \Cbb_f(\lambda)$}, we
call $\Cset[\Xbf] = \Cset \cap \Xbf$ the \emph{empirical density cluster}. The
size of the symmetric set difference between estimated and empirical clusters is 
a commonly used metric to quantify cluster estimation error
\citep{korostelev1993,polonik1995,rigollet2009}.

\begin{definition}[Symmetric set difference]
  \label{def: symmetric_set_diff}
  For an estimator \smash{$\Cest \subseteq \Xbf$} and set
  $\mathcal{S} \subseteq \Reals^d$, we define   
  \begin{equation}
    \label{eqn: misclassification_rate}
    \Delta(\Cest, \mathcal{S}) := \abs{\Cest \setminus \mathcal{S}[\Xbf] \cup
      \mathcal{S}[\Xbf] \setminus \Cest},
  \end{equation}
  the cardinality of the symmetric set difference between 
  \smash{$\Cest$} and $\mathcal{S} \cap \Xbf = \mathcal{S}[\Xbf]$. 
\end{definition}

Note that this symmetric set difference metric does not account for the distance
points in \smash{$\Cest \setminus \mathcal{S}[\Xbf]$} may be from
$\mathcal{S}$  \citep{singh2009}. Hence in this paper, in addition to the symmetric
set difference metric, we also study a second asymptotic notion for the quality
of a cluster estimator introduced by \citet{hartigan1981}.

% which measures whether $\Cest$ can distinguish any two distinct
% elements $\Cset, \Cset' \in \Cbb_f(\lambda)$.

\begin{definition}[Consistent density cluster estimation]
  \label{def: consistent_density_cluster_estimation}
  For an estimator \smash{$\Cest \subseteq \Xbf$} and cluster 
  \smash{$\Cset \in \Cbb_f(\lambda)$}, we say \smash{$\Cest$} is a consistent 
  estimator of $\Cset$ if for all \smash{$\Cset' \in \Cbb_f(\lambda)$} with
  $\Cset \not= \Cset'$, the following holds as $n \to \infty$: 
  \begin{equation}
    \label{eqn: consistent_density_cluster_recovery}
    \Cset[\Xbf] \subseteq \Cest \quad \text{and} \quad
    \Cest \cap \Cset'[\Xbf] = \emptyset,
  \end{equation}
  with probability tending to 1.
\end{definition}

\paragraph{Summary of results.} A summary of our results (and outline for this
paper) is as follows.

\begin{enumerate}
\item In Section \ref{sec: consistent_cluster_estimation_with_ppr}, we introduce
  a set of natural geometric conditions on the density cluster $\Cset$
  %formalize a measure of difficulty based on these geometric conditions, 
  and show that when Algorithm \ref{alg: ppr} is properly initialized, the size of
  the symmetric set difference of \smash{$\Cest$} and a thickened version of the
  density cluster $\Csig$ can be bounded in a meaningful way.
  % this difficulty measure.

\item We further show in Section \ref{sec:
    consistent_cluster_estimation_with_ppr} that if the density cluster 
  $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr}
  will consistently estimate a density cluster in the sense of
  \eqref{eqn: consistent_density_cluster_recovery}.

\item In Section \ref{sec: analysis}, we detail some of the analysis required to
  prove our main results, and expose the parts various geometric quantities play 
  in the difficulty of the clustering problem.

\item In Section \ref{sec: experiments}, we empirically investigate the
  tightness of our analysis, and provide examples showing how violations of our
  geometric conditions impact density cluster recovery by PPR.
\end{enumerate}

Our main takeaway: PPR, run on a neighborhood graph, recovers geometrically
compact high-density clusters.

\paragraph{Related work.} In addition to the background on local spectral
clustering given previously, a few related lines of work are worth
highlighting. \citep{shi2009,schiebinger2015} examine the consistency of  
spectral algorithms in recovering the latent labels in certain
nonparametric mixture models. Their results focus on global rather than local 
methods, and thus impose global rather than local conditions on the nature
of the density. Moreover, they do not in general guarantee recovery of density 
clusters, which is the focus in our work. Perhaps most importantly, these works
rely on general cluster saliency conditions, which implicitly depend on many
distinct geometric aspects of the cluster $\Cset$ under consideration. We make
this dependence more explicit, and in doing so expose the role each geometric
condition plays in the clustering problem.

More broadly, density clustering and level set estimation is a well-studied
problem. \citet{polonik1995, rigollet2009} study density clustering under
symmetric set difference, \citet{tsybakov1997, singh2009} exhibit minimax
optimal level set estimators under Hausdorff loss and \citet{hartigan1981,
  chaudhuri2010} consider consistent estimation of the cluster tree, to note but
a few works in the area. Our goal is not to improve on these results, nor to 
offer a new algorithm for level set estimation; indeed, seen as a density
clustering algorithm, PPR has none of the optimality guarantees of the
methods in the aforementioned works. Rather, our motivation is start with a 
popular  and flexible local spectral algorithm, PPR, and to better understand
and characterize the distinctions between those density clusters which are 
well-conditioned for PPR, and those which are not.

@@

To motivate these conditions, for an arbitrary graph $G = (V,E)$ and subset of vertices $S \subseteq V$, consider the normalized cut $\Phi(S;G)$ (defined as in \eqref{eqn: normalized_cut}), as well as the mixing time of a random walk over the induced subgraph $G[S]$ (defined in Section \ref{sec: analysis} by \eqref{eqn: mixing_time}, and denoted by $\tau_{\infty}(G[S])$). In \citet{zhu2013}, it is shown that for a constant $c > 0$, the \pprspace estimate $\Cest$ of $S$ satisfies
\begin{equation}
\label{eqn: graph_symmetric_set_difference_1}
\vol(\Cest \setminus S; G) + \vol(S \setminus \Cest; G) \leq c \bigl( \Phi(S,G) \cdot \tau_{\infty}(G[S])\bigr) \vol(S;G)
\end{equation}
where the left hand side resembles a (degree-weighted) form of the symmetric set difference of \eqref{eqn: misclassification_rate}.

@@

\begin{theorem}
	\label{thm: misclassification_rate}
	Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and additionally assume $f$ satisfies \ref{asmp: valid_region}. Then, there exists a universal constant $c_4 > 0$ such that with probability tending to $1$ as $n \to \infty$, 
	\begin{equation}
	\label{eqn: misclassification_rate_ub}
	\Delta(\Csig[\Xbf], \Cest) \leq c_4 \kappa(\Cset) \frac{\Lambda_{\sigma}}{\lambda_{\min}}.
	\end{equation}
\end{theorem}

@@

\begin{theorem}
	\label{thm: consistent_recovery_of_density_clusters}
	Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a $\kappa$-well-conditioned cluster (with respect to some $\theta$), and additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm \ref{alg: ppr} is well-initialized, there exists a universal constant $c_5 > 0$ such that if
	\begin{equation}
	\label{eqn: kappa_ub}
	\kappa(\Cset) \leq c_5 \frac{\lambda_{\sigma}^2r^d \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)},
	\end{equation}
	then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}.
\end{theorem}
